Program Equivalence : A Relational Separation Logic interactive prover implemented in Maude

### 2018-07-02 paper/paper_after_guide.tex
replaced black background screenshot with white bg one
printed version of the paper

\lstset{style=mystyle}
\begin{document}
\begin{titlepage}
\begin{center}
\afterpage{\blankpage}
\textbf{
ALEXANDRU IOAN CUZA UNIVERSITY OF IAȘI
}
\\
\textbf{FACULTY OF COMPUTER SCIENCE}
\end{center}
   \vspace{40mm}
\begin{center}
	\Large\textbf {Program Equivalence: An Interactive Relational Separation Logic Prover Implemented in Maude}\\
	\vspace{40mm}
	\large\textit {ANDREI-ALIN CORODESCU}
	\\
	\vspace{20mm}
	\textbf{Session: }\textit{July, 2018}\\
	\vspace{30mm}
	\textbf{Scientific Coordinator}\\
	\textbf{\textit{Conf. Dr. Ciobâcă Ștefan}}
	\vspace{30mm}
\end{center}
\end{titlepage}

@@

The problem this project aims to solve is reducing the gap between theoretical and practical aspects of formal proofs for program equivalence, by creating an executable environment for Relational Separation Logic in which the user can prove the equivalence of programs. Through the creation of such a tool, the theoretical concepts become more accessible, easier to apply and understand. Making formal verification methods more accessible leads to an overall increase in the quality of software, by replacing the existing methods of verifying equivalence, that are often prone to errors, with more robust ones.\\

@@

Each proof rule in the Separation Logic and Relational Separation Logic has been translated into a Maude rewrite rule. The prover is centered around the concept of \textbf{Goals} which can be one of the three things: 
\begin{itemize}
	\item {Hoare Quadruple} - corresponding to Relational Separation Logic;
	\item {Hoare Triple} - corresponding to Separation Logic;
	\item {Implication between relations and assertions}.
\end{itemize} 
The proof rules are interpreted in a bottom-up manner, meaning that, applying a rule to a goal, which must match the conclusion of the used rule, will generate new goals consisting of the hypothesis: Listing \ref{slConseq} presents how a Separation Logic rule has been translated into Maude syntax and Listing \ref{rslConseq} presents the same process for the Relational Separation Logic counterpart of the rule.
\begin{flushleft}
	The Separation Logic rule of \textsf{Consequence}:
\end{flushleft}
	\medskip
	\[
	\begin{prooftree}
	\Hypo{P \Rightarrow P\prime}
	\Hypo{\{P\prime\} C \{Q\prime\}}
	\Hypo{Q\prime \Rightarrow Q}
	\Infer 3 {\{P\}\  C  \{Q\}}
	\end{prooftree}	
	\]
	Is represented in Maude syntax as: 
\begin{lstlisting}[caption=Separation Logic Consequence rule,label=slConseq]
	rl [SeparationLogic-Consequence] : {P} C {Q} => ((P => P1) <> ({P1} C {Q1})) <> (Q1 => Q) [nonexec] .
\end{lstlisting}
\begin{flushleft}
	The Relational Separation Logic rule of \textsf{Consequence}:
\end{flushleft}
	\[
	\begin{prooftree}
	\Hypo{R \Rightarrow R_1}
	\Hypo{\hq{R_1}{C}{C\prime}{S_1}}
	\Hypo{S_1 \Rightarrow S}
	\Infer 3 {\hq{R}{C}{C\prime}{S}}
	\end{prooftree}	
	\]
	is represented in Maude syntax as: 
	\begin{lstlisting}[caption=Relational Separation Logic Consequence rule,label=rslConseq]
	rl [Consequence] : { R } C1 -- C2  { S } => ((R => R1) <> ({R1} C1 -- C2 {S1})) <> (S1 => S) [nonexec] .
	\end{lstlisting}
The \texttt{<>} operator is used to separate sub-goals generated by the rule.
\\
The \texttt{---} operator is used to separate the two snippets of code mentioned in the \textbf{Hoare Quadruple}.
\\

@@

	A sample output of the prover is presented in Figure \ref{fig:CLIScreenShot}, showcasing the manipulation of white spaces (newlines, indentation and spaces), usage of colours to distinguish between different elements - Relations are coloured in green, while the analysed code has its keywords highlighted using cyan. The separator between the two pieces of code is also coloured in red;
	\begin{figure}[h]
		\includegraphics[width=\linewidth]{pictures/CLI.png}
		\caption{Formatted output of the prover}
		\label{fig:CLIScreenShot}
	\end{figure}
	\item The \texttt{metaPrettyPrint} operator in the \texttt{META-LEVEL} module, which returns a term as a list of quoted identifiers to be printed, formatted respecting the \texttt{format} attributes defined for the operators used in the term. An example is presented in Listing \ref{takeFirst}, which presents the rule used to take the top element of the \textbf{GoalStack} for processing, and writes the formatted output of the goal to the \textbf{StagingOutput} field.
	\begin{figure}[h]
	\begin{lstlisting}[label=takeFirst,caption=Rewrite rule making use of metaPrettyPrint]
crl [takeFirstGoalFromList] : < noAction ; RG ; noGoal ; G <> GS ; GL ; nil > 
=> 
< check ; RG ; G ; GS ; GL ; 'current ' 'goal '  '  'is '\n metaPrettyPrint(upModule('PROVER-INTERFACE, false), upTerm(G)) '\n > 
if G =/= noGoal .
	\end{lstlisting}
	\end{figure}
\end{itemize}
All throughout the prover calls to \texttt{metaPrettyPrint} have been made in order to inform the user of what the prover is doing and how.
\\

>>>>

\lstset{style=mystyle}
\begin{document}
\begin{titlepage}
\begin{center}
\textbf{
ALEXANDRU IOAN CUZA UNIVERSITY OF IAȘI
}
\\
\textbf{FACULTY OF COMPUTER SCIENCE}
\end{center}
   \vspace{40mm}
\begin{center}
	\Large\textbf {Program Equivalence: An Interactive Relational Separation Logic Prover Implemented in Maude}\\
	\vspace{40mm}
	\large\textit {ANDREI-ALIN CORODESCU}
	\\
	\vspace{20mm}
	\textbf{Session: }\textit{July, 2018}\\
	\vspace{30mm}
	\textbf{Scientific Coordinator}\\
	\textbf{\textit{Conf. Dr. Ciobâcă Ștefan}}
	\vspace{30mm}
\end{center}
\end{titlepage}

@@

The problem this project aims to solve is program equivalence, by reducing the gap between theoretical and practical aspects of formal proofs for it. The project creates an executable environment for Relational Separation Logic in which the user can prove the equivalence of programs. Through the creation of this tool, the theoretical concepts become more accessible, easier to apply and understand. Making formal verification methods more accessible leads to an overall increase in the quality of software, by replacing the existing methods of verifying equivalence, that are often prone to errors, with more robust ones.\\

@@

Each proof rule in the Separation Logic and Relational Separation Logic has been translated into a Maude rewrite rule. The prover is centered around the concept of \textbf{Goals} which can be one of the three things: 
\begin{itemize}
	\item {Hoare Quadruple} - corresponding to Relational Separation Logic;
	\item {Hoare Triple} - corresponding to Separation Logic;
	\item {Implication between relations and between assertions}.
\end{itemize} 
The proof rules are interpreted in a bottom-up manner, meaning that, applying a rule to a goal, which must match the conclusion of the used rule, will generate new goals consisting of the hypothesis: Listing \ref{slConseq} presents how a Separation Logic rule has been translated into Maude syntax and Listing \ref{rslConseq} presents the same process for the Relational Separation Logic counterpart of the rule.
\begin{flushleft}
	The Separation Logic rule of \textsf{Consequence}:
\end{flushleft}
	\medskip
	\[
	\begin{prooftree}
	\Hypo{P \Rightarrow P\prime}
	\Hypo{\{P\prime\} C \{Q\prime\}}
	\Hypo{Q\prime \Rightarrow Q}
	\Infer 3 {\{P\}\  C  \{Q\}}
	\end{prooftree}	
	\]
	Is represented in Maude syntax as: 
\begin{lstlisting}[caption=Separation Logic Consequence rule,label=slConseq]
	rl [SeparationLogic-Consequence] : {P} C {Q} => ((P => P1) <> ({P1} C {Q1})) <> (Q1 => Q) [nonexec] .
\end{lstlisting}
\begin{flushleft}
	The Relational Separation Logic rule of \textsf{Consequence}:
\end{flushleft}
	\[
	\begin{prooftree}
	\Hypo{R \Rightarrow R_1}
	\Hypo{\hq{R_1}{C}{C\prime}{S_1}}
	\Hypo{S_1 \Rightarrow S}
	\Infer 3 {\hq{R}{C}{C\prime}{S}}
	\end{prooftree}	
	\]
	is represented in Maude syntax as: 
	\begin{lstlisting}[caption=Relational Separation Logic Consequence rule,label=rslConseq]
	rl [Consequence] : { R } C1 -- C2  { S } => ((R => R1) <> ({R1} C1 -- C2 {S1})) <> (S1 => S) [nonexec] .
	\end{lstlisting}
The \texttt{<>} operator is used to separate sub-goals generated by the rule.
\\
The \texttt{---} operator is used to separate the two snippets of code mentioned in the \textbf{Hoare Quadruple}.
\\

@@

	A sample output of the prover is presented in Figure \ref{fig:CLIScreenShot}, showcasing the manipulation of white spaces (newlines, indentation and spaces), usage of colours to distinguish between different elements - Relations are coloured in green, while the analysed code has its keywords highlighted using cyan. The separator between the two pieces of code is also coloured in red;
	\begin{figure}[h]
		\includegraphics[width=\linewidth]{pictures/bg_white.png}
		\caption{Formatted output of the prover}
		\label{fig:CLIScreenShot}
	\end{figure}
	\item The \texttt{metaPrettyPrint} operator in the \texttt{META-LEVEL} module, which returns a term as a list of quoted identifiers to be printed, formatted respecting the \texttt{format} attributes defined for the operators used in the term. An example is presented in Listing \ref{takeFirst}, which presents the rule used to take the top element of the \textbf{GoalStack} for processing, and writes the formatted output of the goal to the \textbf{StagingOutput} field.
	\begin{figure}[h]
	\begin{lstlisting}[label=takeFirst,caption=Rewrite rule making use of metaPrettyPrint]
crl [takeFirstGoalFromList] : < noAction ; RG ; noGoal ; G <> GS ; GL ; nil > 
=> 
< check ; RG ; G ; GS ; GL ; 'current ' 'goal '  '  'is '\n metaPrettyPrint(upModule('PROVER-INTERFACE, false), upTerm(G)) '\n > 
if G =/= noGoal .
	\end{lstlisting}
	\end{figure}
\end{itemize}
All throughout the prover calls to \texttt{metaPrettyPrint} have been made in order to inform the user of what the prover is doing and how.
\\
